\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-23 pldi/draft.tex
Pass on section 4

\heading{Edges---Streams}
Edges in the DFG represent streams, the basic data abstraction of the shell.
They are used as communication channels between nodes in the graph, and as the input or output of the entire graph.
Edges are represented as possibly unbounded streams of type $D\kstar$.
Edges can either refer to named files or FIFO pipes used for interprocess communication.
Edges that do not start from a node in the graph represent the graph input;
  edges that do not point to a node in the graph are its outputs.
\nv{OK This distinction is explained in this par; I would put this earlier}

@@

\heading{\todo{Streaming} Commands}
These commands consume their inputs sequentially and one
element at a time, with the possible exception of static input files,
and produce one output file. The quintessential example of
\todo{streaming} commands is \ttt{cat}, that consumes its inputs in
order, producing their concatenation as output. A more interesting
example is \ttt{comm -23 f1 f2}, which has \ttt{f2} as its static
input (it needs to read it completely before consuming anything from
\ttt{f1} and then consumes \ttt{f1} one element at a time. These
commands can be represented as functions of type $f : D\kstar \times
[D\kstar] \to D\kstar$, where the first argument represents the
concatenation of the inputs that are consumed one at a time and the
second argument represents the static inputs.
%% \km{Again, a graphical
%%   presentation of these operators would be useful. Maybe with a way to
%%   distinguish between between streaming inputs and static inputs? Btw,
%%   this distinction is not completely clear to me at this
%%   point. Clearly, not all programs that we can write are streaming
%%   (monotone).} \kk{Do you mean that there are shell scripts that are
%%   non monotone? That is true but we are assuming monotone nodes. Do I
%%   miss something?}
We focus on these commands because a large subset of \sta and \pur
commands fall in this category.
\nv{Based on the intro of 4 or 4.1 I did not expect streaming. I understand edges and nodes, but how do streaming commands fit into the picture? the answer might be the last sentence which should be first---reviewer: why am I reading what I'm reading?:P}

@@

\heading{Stateless and Parallelizable Pure Commands}
As mentioned in \Cref{cmd}, stateless commands such as \ttt{tr}
operate independently on individual elements of the stream
(characters, lines, or files) without maintaining any state. In this
work, we consider lines as the data quantum, so only commands that are
stateless regarding lines (or any finer granularity such as
characters) are considered stateless. Formally, a
\todo{streaming} command $f$ is stateless if it commutes with
the operation of concatenation, \ie it is a semigroup homomorphism:
%% with respect to concatenation \km{It commutes with the operation of concatenation. This is the definition of (semigroup) homomorphism. I am not sure that linearity is an appropriate term.}
%% , \ie satisfying the following equation:

@@

\heading{Parallelization Transformations}
Based on these equations, we can define a \emph{node parallelization
  transformation} on a stateless node $v$ that is preceded by a
\ttt{cat} with $n$ input streams and is followed by a node $v'$ (Fig.~\ref{fig:parallelization-transformation}). The
transformation replaces $v$ with $n$ new nodes, routing each of the
$n$ input streams to one of them, and commutes the \ttt{cat} node
after them to concatenate their outputs and transfer them to
$v'$. Since each incoming edge represents a stream of data $x_i :
D^*$, and the only behaviour of a DFG is its output, this
optimization $v(x_1 \cdot x_2 \cdots x_n, s) \Rightarrow v(x_1, s)
\cdot v(x_2, s) \cdots v(x_n, s)$ can be shown to preserve the
behavior of the graph.

@@

\heading{Auxiliary Transformations}
\todo{We} also define a set of auxiliary transformations that are shown in
\Cref{fig:auxiliary-transformations}.  The first two \nv{Name them?} enable the
parallelization transformations by inserting \ttt{cat} nodes,\footnote{\nv{This is minor, but for the model maybe name them concatenation nodes or ++? rather than \ttt{cat}}} either
by using them to concatenate many inputs of one command, or by first
adding a \ttt{split} node that splits the input in batches. The final
one inserts a relay that performs the identity transformation on its
input. Relay nodes can be useful for monitoring and debugging, as well
as for performance improvements~\sx{optimizer}.

>>>>

\heading{Edges---Streams}
Edges in the DFG represent streams, the basic data abstraction of the shell.
They are used as communication channels between nodes in the graph, and as the input or output of the entire graph.
Edges are represented as possibly unbounded streams of type $D\kstar$.
Edges can either refer to named files or FIFO pipes used for interprocess communication.
Edges that do not start from a node in the graph represent the graph input;
  edges that do not point to a node in the graph are its outputs.
%% \nv{OK This distinction is explained in this par; I would put this earlier}

@@

\heading{Streaming Commands}
%
A large subset of the parallelizable \tsta and \tpur classes falls
into the special category of streaming commands. These commands
consume their inputs sequentially and one element at a time, with the
possible exception of static input files, and produce one output
file. The quintessential example of streaming commands is \ttt{cat},
that consumes its inputs in order, producing their concatenation as
output. A more interesting example is \ttt{comm -23 f1 f2}, which has
\ttt{f2} as its static input (it needs to read it completely before
consuming anything from \ttt{f1} and then consumes \ttt{f1} one
element at a time. These commands can be represented as functions of
type $f : D\kstar \times [D\kstar] \to D\kstar$, where the first
argument represents the concatenation of the inputs that are consumed
one at a time and the second argument represents the static inputs.
%% \km{Again, a graphical
%%   presentation of these operators would be useful. Maybe with a way to
%%   distinguish between between streaming inputs and static inputs? Btw,
%%   this distinction is not completely clear to me at this
%%   point. Clearly, not all programs that we can write are streaming
%%   (monotone).} \kk{Do you mean that there are shell scripts that are
%%   non monotone? That is true but we are assuming monotone nodes. Do I
%%   miss something?}
%% We focus on these commands because a large subset of \sta and \pur
%% commands fall in this category.
%% \nv{Based on the intro of 4 or 4.1 I did not expect streaming. I understand edges and nodes, but how do streaming commands fit into the picture? the answer might be the last sentence which should be first---reviewer: why am I reading what I'm reading?:P}

@@

\heading{Stateless and Parallelizable Pure Commands}
As mentioned in \Cref{cmd}, stateless commands such as \ttt{tr}
operate independently on individual elements of the stream
(characters, lines, or files) without maintaining any state. In this
work, we consider lines as the data quantum, so only commands that are
stateless regarding lines (or any finer granularity such as
characters) are considered stateless. Formally, a
streaming command $f$ is stateless if it commutes with
the operation of concatenation, \ie it is a semigroup homomorphism:
%% with respect to concatenation \km{It commutes with the operation of concatenation. This is the definition of (semigroup) homomorphism. I am not sure that linearity is an appropriate term.}
%% , \ie satisfying the following equation:

@@

\heading{Parallelization Transformations}
%
Based on these equations, we can define a \emph{node parallelization
  transformation} on a stateless node $v$ that is preceded by a
concatenation, i.e. the command \ttt{cat}, of $n$ input streams and is
followed by a node $v'$
(Fig.~\ref{fig:parallelization-transformation}). The transformation
replaces $v$ with $n$ new nodes, routing each of the $n$ input streams
to one of them, and commutes the \ttt{cat} node after them to
concatenate their outputs and transfer them to $v'$. Since each
incoming edge represents a stream of data $x_i : D^*$, and the only
behaviour of a DFG is its output, this optimization $v(x_1 \cdot x_2
\cdots x_n, s) \Rightarrow v(x_1, s) \cdot v(x_2, s) \cdots v(x_n, s)$
can be shown to preserve the behavior of the graph.

@@

\heading{Auxiliary Transformations}
%
\sys also performs a set of auxiliary transformations that are
depicted in \Cref{fig:auxiliary-transformations}. If a node has many
inputs, \sys concatenates these inputs by inserting a \ttt{cat} node
to enable the parallelization transformations. In cases where a
parallelizable node has one input and is not preceded by a
concatenation, \sys inserts a \ttt{cat} node that is preceded by its
inverse \ttt{split}, so that the concatenation can be commuted with
the node. The third depicted transformation inserts a relay node that
performs the identity transformation. Relay nodes can be useful for
monitoring and debugging, as well as for performance
improvements~\sx{optimizer}.

%% The first two
%% \nv{Name them?}  enable the parallelization transformations by
%% inserting \ttt{cat} nodes,\footnote{\nv{This is minor, but for the
%%     model maybe name them concatenation nodes or ++? rather than
%%     \ttt{cat}}} either by using them to concatenate many inputs of one
%% command, or by first adding a \ttt{split} node that splits the input
%% in batches. The final one inserts a relay that performs the identity
%% transformation on its input.
